---
title: 'EDS 241: Assignment 4'
author: "Alexandra Yousefivand"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  \setlength{\parindent}{1em}
  \usepackage{float}
  \renewcommand{\thesection}{Question (\alph{section})}
---


``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3,
                      echo = TRUE, message = FALSE, warning = FALSE)                     
```


```{r}
# load packages
packages=c("stargazer", "tidyverse", "estimatr", "janitor", "car", "ivreg")
#           "dplyr","stringr", , "cowplot",
#           "ggplot2", "tinytex", "datasets", "tibble"

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

#devtools::install_github('rstudio/rmarkdown')
options(scipen=999) # not scientific notation
```


```{r load data}
# load data
data_raw <- read.csv("EU_sardines.csv")

data <- data_raw %>% clean_names()
```


# (a)

_Estimate a bivariate regression of log(volume_sold_kg) on log(price euro_kg). What is the price elasticity of demand for sardines? Test the null hypothesis that the price elasticity is equal to -1._


```{r}
data_log <- data %>% mutate(vol_log = log(volume_sold_kg),
                            price_log = log(price_euro_kg))
```



```{r bivariate regression}
model <- lm(formula = volume_sold_kg ~ price_euro_kg,
            data = data_log,
            family = binomial(link = "logit"))
summary(model)
```
```{r}
confidence_interval <- confint(model)
conf_int_low <- confidence_interval[[2]]
conf_int_high <- confidence_interval[[4]]
```


```{r}
coeff_a <- round(model$coefficient[[2]], 2)
coeff_a
```



**The estimated price elasticity of demand for sardines is `r coeff_a`. There is 95% probability that the interval [`r round(conf_int_low, 2)`, `r round(conf_int_high, 2)`] contains the true price elasticity of demand for sardines. Because this confidence interval does not contain the value -1, we reject the null hypothesis that the price elasticity is equal to -1.**




# (b)

_Like in Lecture 8 (see the IV.R script), we will use wind_m_s as an instrument for
log(price_euro_kg). To begin, estimate the first-stage regression relating log(price_euro_kg) to wind_m_s. Interpret the estimated coefficient on wind speed. Does it have the expected sign? Also test for the relevance of the instrument and whether it is a “weak” instrument by reporting the proper F-statistic._


```{r first stage regression}
model_b <- lm_robust(formula = price_log ~ wind_m_s, data = data_log)
summary(model_b)
```

```{r}
wind_coeff <- round(model_b$coefficients[[2]], 5)
```


**The estimated coefficient on wind speed**:
for each 1 unit (m/s) increase in wind speed, there is `r wind_coeff` increase (+) in the log() of the price (euro/kg).

This positive coefficient matches intuition, as this indicates: as wind increases, conditions for fishing become more difficult; fewer sardines are caught (supply down), and therefore the price increases (demand up).


```{r F-stat}
f_stat_model_b <- linearHypothesis(model_b, c("wind_m_s = 0"),
                                 white.adjust = "hc2")
f_stat_b <- f_stat_model_b$F[[2]]
f_stat_b # don't know why this won't knit
```

**The F-statistic is `r f_stat_b`. This value is greater than 10, indicating wind speed is a relevant and non-weak instrument.**


# (c)


_Estimate the TSLS estimator of the price elasticity of demand for sardines using wind_m_s as an instrument for log(price_euro_kg). What is the estimated price elasticity of demand for sardines?_



```{r two-stage least squares}
tsls_c <- ivreg(formula = vol_log ~ price_log | wind_m_s, 
              data = data_log)
summary(tsls_c)
```

```{r}
instrument_wind_coeff <- round(tsls_c$coefficients[[2]], 3)
instrument_wind_coeff
```


**The estimated price elasticity of demand for sardines is `r instrument_wind_coeff`. In other words, for every one euro increase in log price, there is a `r instrument_wind_coeff` unit increase in volume**.

# (d)

_Repeat the exercise in (c), but include fixed effects for each year, month, and country. [Hint: you can use the command “as.factor(country) + as.factor(year) +as.factor(month)” to the ivreg function in R. Report the estimated price elasticity of demand and the F-statistic testing for relevant and non-weak instruments._



```{r}
tsls_d <- ivreg(data = data_log,
                formula = vol_log ~ price_log
                                       + as.factor(country)
                                       + as.factor(year)
                                       + as.factor(month)
                                     | wind_m_s
                                       + as.factor(country)
                                       + as.factor(year)
                                       + as.factor(month))
summary(tsls_d)
```


```{r}
tsls_wind_coeff <- round(tsls_d$coefficients[[2]], 3)
tsls_wind_coeff
```

**The estimated price elasticity of demand is `r tsls_wind_coeff`.** 


```{r}
model_d <- lm_robust(data = data_log,
              formula = price_log ~ wind_m_s
                                    + as.factor(country)
                                    + as.factor(year)
                                    + as.factor(month))
summary(model_d)
```




```{r}
f_stat_model_d <- linearHypothesis(model_d, c("wind_m_s = 0"),
                     white.adjust = "hc2")
f_stat_d <- round(f_stat_model_d$Chisq[[2]], digits = 3)
f_stat_d
```

**The F-statistic is `r f_stat_d`. This value is greater than 10, indicating wind speed is a relevant and non-weak instrument.**
